---
author: "Gergana Yordanova"
date: "2/15/2022"
title: "Lab6"
output: html_document
---

install.packages("nycflights13")

```{r setup, include = FALSE}
library(nycflights13)   
library(tidyverse)
```

## Part 1

### Question 1
```{r, echo=FALSE}
flights %>% select(flight, carrier, dep_time, year, month, day) 
```

### Question 2
```{r, echo=FALSE}
flights %>% select(dep_time,dep_delay,arr_time,arr_delay)
```

### Question 3
```{r, echo=FALSE}
flights %>% select(contains("arr_"))
```

### Question 4
The contain function gets the arguments that start with arr_ and prints them out, unlike select where one says what arguments they want printed

## Part 2

### Question 5
```{r, echo=FALSE}
?flights
flights %>% filter(arr_delay >= 120)
```

### Question 6 
### a 
```{r, echo=FALSE}
filter(flights,dest == "IAH" | dest == "HOU")
```

### b
```{r, echo=FALSE}
filter(flights,dep_delay > 60 & arr_delay < 30)
```

### Question 7 
```{r, echo=FALSE}
flights %>% filter(is.na(dep_time))
```

### Question 8
Other variables that are missing data are the departures delays and arrival times. I think this data is missing because there is no departure time to begin with. So since there was no actual departure, there is no departure delay or arrival time. 


## Part 2
### Question 9
```{r, echo=FALSE}
flights %>% arrange(dep_time)
```

### Question 10 
### a
```{r, echo=FALSE}
flights %>% arrange(desc(dep_delay))
```

### b
```{r, echo=FALSE}
flights %>% arrange(distance)
```


## Part 4 ?
### Question 11
```{r, echo=FALSE}
flights %>% count(tailnum) %>% arrange(desc(n))
```

### Question 12
```{r, echo=FALSE}
flights %>% filter((month==6 & day>=21) | (month == 9 & day<=21)| (month==7)|(month==8)) %>% count(month,day)
```

### Question 13
```{r, echo=FALSE}
flights %>%  count(distance>100 & distance<200)
```

### Question 14
```{r, echo=FALSE}
flights %>% filter(air_time >420) %>% count(dest)
```







