---
author: "Gergana Yordanova"
date: "2/22/2022"
title: "Lab7"
output: html_document
---

## Part 1

install.packages("nycflights13")

```{r setup, include = FALSE}
library(nycflights13)   
library(tidyverse)
```

```{r}
flights %>% select(year:day, ends_with("delay"), distance, 
air_time) %>% mutate(hours = air_time / 60)
```

### Question 1
mutate() puts the new hours column at the very end of the data table on the right hand side. 

### Question 2
```{r, echo=FALSE}
flights %>% select(air_time, arr_time, dep_time)  %>% 
  mutate(time_diff=arr_time-dep_time)
```

### Question 3
I see how the air time describes how long the plain was in the air. The time_diff variable represents the difference between arrival time and departure time. The time_diff is longer than the air time.

### Question 4 
```{r, echo=FALSE}
min_since_midnight <- function(x) {
	if ((x %% 100) > 59 || (x - x %% 100) > 2300) {
		return(NA)
	} else {
		return ((x - x %% 100)/100 * 60 + x %% 100)
	}
}

flights %>% select(air_time, arr_time, dep_time)  %>% 
  mutate(dep_time_min=min_since_midnight(dep_time),
         arr_time_min=min_since_midnight(arr_time))
```
### Question 5
I think the output should be a bigger number since it calculates the minutes. I believe it did the right thing because when I input the dep_time for 517 (am) and calculated the minutes from 12:00 AM, I got 317 minutes.This is the number that I got for dep_time_min for that particular time. 


## Part 2
### Question 6
```{r, echo=FALSE}
flights %>%  
   group_by(carrier) %>%  
   summarize(avg_delay = mean(dep_delay, na.rm = TRUE)) 
```
### Question 7
```{r, echo=FALSE}
flights %>%  
   group_by(carrier) %>%  
   summarize(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  arrange(desc(avg_delay))
```
The carrier that had the worst delay is F9

### Question 8
```{r, echo=FALSE}
flights %>%  
   group_by(carrier) %>%  
   summarize(avg_delay = mean(dep_delay, na.rm = TRUE),
             num_flights= n()) %>% 
  arrange(desc(avg_delay)) 
```

### Question 9 
```{r, echo=FALSE}
flights %>% group_by(dest) %>% 
  summarize(arrival_delay = sum(arr_delay, na.rm = TRUE))
```


### Question 10 
```{r, echo=FALSE}
flights %>% 
  filter(arr_delay==10) %>%
  group_by(flight) %>%
  summarize(num_flights=n())
```





