---
author: "Gergana Yordanova"
date: "3/15/2022"
title: "Lab9"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
#install.packages("gapminder")
library(gapminder)
?gapminder
jobs <- read_csv("jobs.csv")
t_passengers <- read_csv("Titanic_Passengers.csv")
```

## Part 1
### Question 1
The gapminder data has 1704 rows and 6 variables. The variables consist of country, continent, year, life expectancy, population, and GDP per capita. A single observation is about life expectancy, GDP per capita and the popualation by country. The source is from http://www.gapminder.org/data/. 

The jobs dataset has 558 rows and 1 column. The variables consist of year, sex, occupation, people, total number of people for that year, and proportion. A single observation is about reported occupations in the United States labor force from 1850-2000. The source is from 1950. 

The titanic passengers has 558 rows and 1 column. The variables consist of passenger id, survived, passenger class, first name, last name, sex, age, number of siblings,number of parents, ticket number, passenger fare, cabin, and port of Embarkation. A single observation is about the kind of passenger and its characteristics they had on the Titanic. 

### Question 2
```{r}
jobs %>% 
  filter(year>=1910 & year<=2000) %>%
  filter(occupation == "Clerical Worker")%>% 
   summarize(mean_jobs = mean(proportion, na.rm = TRUE),
             std = sd(proportion, na.rm = TRUE))
```

### Question 3 
```{r}
jobs %>%
  group_by(occupation) %>%
  summarize(total = n_distinct(occupation),
            total = n())
```
### Question 4 
```{r}
jobs %>%
  group_by(occupation) %>%
  summarize(max_proportion = max(proportion))%>% 
  arrange(desc(max_proportion))
```
Farmer had the highest maximum proportion. 
## Part 2


### Question 5
```{r}
t_passengers %>%
  ggplot(aes(Age)) + geom_histogram()
```

### Question 6
```{r}
t_passengers %>%
  ggplot(aes(Age)) + geom_histogram(bins=50)
```

### Question 7
```{r}
t_passengers %>%
  ggplot(aes(Age)) + geom_histogram(binwidth=5)
```

### Question 8

The appearance in question 5, the graph is with the default parameters to create the histogram. The graph in question 6 has altered the bins which means there will be more bars representing the data. In question 7, the binwidth has changed so the width of the bars is thicker. 

### Question 9
```{r}
t_passengers %>%
  group_by(Sex) %>%
  summarize(avg_sex = mean(Age, na.rm=T))
```
### Question 10 
```{r}
t_passengers %>%
  group_by(Pclass) %>%
  summarize(avg_fare = mean(Fare, na.rm=T),
            std = sd(Fare, na.rm = TRUE),
            total_pop = sum(Fare))
```

### Question 11
```{r}
t_passengers %>%
  group_by(Pclass) %>%
  summarize(max_fare = max(Fare))
```
Class 1 had the most expensive fare with a cost of 512.3292.


```{r} 
t_passengers %>%
  group_by(Pclass) %>%
  summarize(max_variability = sd(Fare)) 
```
Class 1 had the highest variability with standard deviation of 78.38037.

```{r} 
t_passengers %>%
  group_by(Pclass) %>%
  summarize(max_passenger = max(PassengerId)) 
```
Class 3 had the most passengers with 891 people. 

## Part 2

### Question 12
```{r}
gapminder %>% 
  filter(continent=="Europe") %>% 
  ggplot(aes(x=fct_rev(country), y=gdpPercap)) + 
    geom_boxplot() + 
    coord_flip()
```

### Question 13
```{r}
gapminder %>% 
  filter(continent=="Europe") %>% 
  ggplot(aes(x=fct_rev(country), y=gdpPercap)) + 
    geom_boxplot() + 
    coord_flip()
```

My answer from question 12 is the same for question 13


### Question 14
```{r}
gapminder %>% 
  filter(country %in% c("Hungary", "Iceland", 
                        "Greece", "Switzerland")) %>% 
  ggplot(aes(x=fct_rev(country), y=gdpPercap)) + 
    geom_boxplot() + 
    coord_flip()
```

### Question 15
The line in the middle of the box plot represents the median.








